---
title: "Assignment - Week 2"
author: "Reider Ladislau"
date: "April 14, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. Loading and preprocessing the data
``` {r loaddata}
options(scipen = 3)
setwd("E:/Coursera/Reproducible Research/Week 2/Assignment")
activity <- read.csv(file = "./activity.csv", sep = ",", stringsAsFactors = FALSE, header = TRUE)
# Remove NAs
data_proc <- activity[complete.cases(activity), ]
```


## 2. Create Histogram with the total number of steps taken each day
``` {r nrsteptperday}
library(dplyr)
library(magrittr)
steps_per_day <- data_proc %>% dplyr::group_by(date) %>% 
        dplyr::summarise(total = sum(steps))
mean_value   <- round(mean(steps_per_day$total, na.rm = TRUE), digits = 1)
median_value <- round(median(steps_per_day$total, na.rm = TRUE), digits = 1)
```


Create the histogram with the number of steps taken each day:
``` {r createhist}
hist(steps_per_day$total, main = "Total number of steps per day", 
     xlab = "Total number of steps per day", col = "red")
```


## 3. Add Mean and median number of steps taken each day

Mean of total number of steps per day is `r mean_value`, median is `r median_value`.


## 4. Add Time series plot of the average number of steps taken
``` {r createtimeseriesplot}
avg_steps <- aggregate(steps ~ interval, data_proc, mean)

plot(as.numeric(avg_steps$interval), avg_steps$steps, type = 'l', 
     main = "Average number of steps per day", xlab = "Interval", 
     ylab = "Average number of steps", col = "red")
```


## 5. Add The 5-minute interval that, on average, contains the maximum number of steps
``` {r maximumsteps}
# Row which has the highest average steps
row_nr <- which.max(avg_steps$steps)

# Identify interval and the average steps for this interval
max_interval <- avg_steps[row_nr, ]$interval
max_steps <- round(avg_steps[row_nr, ]$steps, digits = 1)
```

The 5-min interval that, on average contains the maximum number of steps is `r max_interval` and the number of steps for this interval is  `r max_steps`


## 6. Add code to describe and show a strategy for imputing missing data
``` {r missingdata}
# Number of rows with missing data
sum(is.na(activity))

# The adopted strategy is to replace the missing values with the average steps in that interval across all the days
for (i in 1:nrow(activity)) {
    if(is.na(activity$steps[i])) 
        {
        value <- avg_steps$steps[which(avg_steps$interval==activity$interval[i])]
        activity$steps[i] <- value 
        }
}

steps_per_day_imputed <- aggregate(steps ~ date, activity, sum)
```


## 7. Create histogram of the total number of steps taken each day after missing values are imputed

``` {r histplotimputed}
# Create histogram plot
hist(steps_per_day_imputed$steps, main = "Total number of steps per day imputed", xlab = "Total number of steps per day", col = "red")
```


Compare the mean and median with and without NAs
``` {r comparison}
# get average and median of imputed data
mean_imputed   <- round(mean(steps_per_day_imputed$steps), digits = 1)
median_imputed <- round(median(steps_per_day_imputed$steps), digits = 1)
```

The mean value without imputed data is `r mean_value`
The mean value with imputed data is `r mean_imputed`
The median value without imputed data is `r median_value`
The median value with imputed data is `r median_imputed`
As a conseuence the mean values do not change, but there is a very slight difference in the meadian values.


## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

``` {r weekday_weekend}
library(ggplot2)
# Create extra column with Weekday or Weekend
activity$Day <- weekdays(as.Date(activity$date))
activity$Day[activity$Day  %in% c('Saturday','Sunday') ] <- "Weekend"
activity$Day[activity$Day != "Weekend"] <- "Weekday"

# Convert type from character to factor
activity$Day <- as.factor(activity$Day)

# Calculate average steps by interval
avg_steps_per_day_imputed <- aggregate(steps ~ interval + Day, activity, mean)

# Create plot
ggplot2::ggplot(avg_steps_per_day_imputed, aes(interval, steps)) +
    geom_line(stat = "identity", aes(colour = Day)) +
    theme_gray() +
    facet_grid(Day ~ ., scales = "fixed", space = "fixed") +
    labs(x = "Interval", y = expression("Number of Steps")) +
    ggtitle("Number of steps per Interval (Weekend or Weekday)")
```

